---
title: "Working with the Kulan package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with the Kulan package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(Kulan)
```

## Key functions

### Drone flight descriptors

**Import some data**

The package has some flight datasets inside it that can be used for generating examples. You can add your own data using any of the read functions available in  http://www.sthda.com/english/wiki/importing-data-into-r. Usually we use the read.csv() to read in a .csv file. There are options for excel files and text files too. 


```{r get the data}
df<-Kulan::All_Flight_Right
```

**Tidy the data frame to make it more R friendly**

The dataframe has some column names that are not very "R friendly" because, for example, they have spaces in the names (e.g. "Alt above ground" ). The {janitor} package replaces the names with a more standard format ("alt_above_ground")

```{r use janitor to tidy the coloumn names }
df<-df %>% 
  janitor::clean_names()

```

**Get the ground in the image**

To calculate the width of a drone image we need to know the height of the drone and the angle of the cameras. The inbuilt functions in the {Kulan} package use the camera dimensions from the cameras used in the field as default values. 

```{r ground in image}
df<-df %>%
  rowwise() %>%
  mutate("grd_in_img"=Kulan::grd_in_img(`alt_above_ground`)) 

mean(df$grd_in_img)
```

**Get the ground surface distance**

To calculate the ground surface distance (1 pixel) there are two functions, one that assumes no distortion due to the angle of the camera and one that allows for distortion (this later function is still in development) . 

```{r}

df<-df %>%
  rowwise() %>%
  mutate("get_GSD"=Kulan::get_GSD(Altitude = `alt_above_ground`)) 

mean(df$get_GSD, na.rm=TRUE) # this includes the drones take off and landing phases so needs to be recalculated 
```


### Abundance estimators

**Simulate a dataset**

To simulate a dataset we can use the sim_kulan_data() function. To get help from the package please use the code below:

```{r, eval=FALSE}
help("sim_kulan_data")
```

which outputs the details of the function:

--- 

### sim_kulan_data {Kulan}	R Documentation

### Simulate a dataset

#### Description

Simulate a dataset

#### Usage

sim_kulan_data(n = 100, ma = 25, p0 = 0.5, ml = 20, w = 0.2)

#### Arguments

n	number of transects

ma	mean number of animals counted where they are counted

p0 proportion of the transects that will be empty

ml	mean length of the transects

w	width of the transects (assuming a constant width)

Value simulated data in a dataframe

---


```{r}
data=Kulan::sim_kulan_data()

```


**Run the Jolly2 model**

To run the Jolly2 model we can use the jolly2() function. Again using help() will output a description of the function.

```{r, eval=FALSE}
help("jolly2")
```

---

### jolly2 {Kulan}	R Documentation

### Calculate Jolly II estimate and 95% Confidence Interval

#### Description

Calculate Jolly II estimate and 95% Confidence Interval

#### Usage

jolly2(species_count, Transect_area, Z = Z)

#### Arguments
species_count	the column from the dataset with the species counts in

Transect_area	the column from the dataset with the transect areas in

Z	the stratum area (the total area that we are extrapolating to)

Value dataframe with Jolly II estimate and 95% CI

---



```{r}
jolly2(species_count = data$sp_count, Transect_area = data$Trans_area,Z=164*5)
```


**Run the n-mixture model**

Finally, to run the n-mixture model you can use the run_kulan_model() function. See more details about this function using help(). 

```{r, eval=FALSE}
help("run_kulan_model")
```


---

### run_kulan_model {Kulan}	R Documentation

### Calculate Bayesian estimate and 95% HDI

#### Description

Calculate Bayesian estimate and 95% HDI

#### Usage

run_kulan_model(
  p_area = p_area.,
  p_min = p_min.,
  p_max = p_max.,
  no_ind = no_ind.,
  plot.out = "No"
)

#### Arguments

p_area	proportion of the stratum that has been sampled

p_min	prior minimum number of animals

p_max	prior maximum number of animals

no_ind Indivduals count column from data

plot.out Add a plot ("Yes") of the distribution of the estimate (default is "No")

Value dataframe with estimate and hdi

---

```{r}
mod1<-run_kulan_model(p_area = .49, p_min = 1000, p_max=4000, no_ind = data$sp_count, plot.out = TRUE)
```


